% !TEX root = main.tex
%\myparagraph{Key points}
%\begin{enumerate}[1.]
%%	\item	Session $\pi$ calculus with process passing. DONE
%%	\item	Identify session $\pi$ and process passing subcalculi and their polyadic variants. DONE
%%	\item	Bisimulation theory for higher-order session semantics. DONE
%%	\item	New triggered bisimulation, related to J\&R's. DONE
%%	\item   Elementary values key to characterizations of behavioural equivalence. DONE
%	\item	Types provide techniques to prove completeness without matching. \jp{TBD}
%	\item	We are interested in encodings with properties a la Gorla. 
%                We extended them to typed setting. \jp{TBD}
%%	\item	Encode name-passing to pure process abstraction calculus, with name abstractions. DONE
%%	\item	Type of the recursion encoding uses non tail recursive type $\trec{t}{\btinp{t} \tinact}$. DONE
%%	\item	Encode higher-order semantics to first order semantics. DONE
%%	\item	Negative result. Cannot encode shared names using only shared names.
%%	\item   Extensions with higher-order abstractions and polyadicity also explored. DONE
%\end{enumerate}

%\smallskip 
%
%\myparagraph{Important things to explain}
%Explain our \HO is very small without containg name passing 
%\[ 
%\abs{x}.P \quad \appl{x}{u}
%\]

%Explain we input only characteristic processes.  
%
%\[
%\lambda x.\mapchar{S}{x}
%\]

%\subsection{Higher-Order Session Calculi}
%\noindent
\paragraph{Context.}
In \emph{higher-order process calculi} 
communicated
values %exchanged in communications 
may contain  processes.
Higher-order concurrency has received significant attention 
from untyped and typed perspectives; see, e.g.,~\cite{ThomsenB:plachoasgcfhop,SangiorgiD:expmpa,JeffreyR05,DBLP:journals/iandc/LanesePSS11,DBLP:journals/cl/KoutavasH12,MostrousY15}.
%=== One alternative
%The combination of features from the $\lambda$-calculus and the $\pi$-calculus
%enables \emph{higher-order process calculi} to exchange values that may contain  processes.
%=== Dimitris version: 
%The combination of features from the $\lambda$-calculus and the $\pi$-calculus,
%in \emph{higher-order process calculi} allows for exchanged values to contain  processes. 
%=== Previous version: 
%By combining features from the $\lambda$-calculus and the $\pi$-calculus, 
%in \emph{higher-order process calculi} exchanged values may contain  processes. 
In this work, we consider \HOp, a higher-order process calculus with \emph{session communication}:
it 
combines functional constructs (abstractions/applications, as in the call-by-value $\lambda$-calculus)
and 
concurrent primitives (synchronisation on shared names, 
communication on linear names, 
  %(value passing, labelled choice), 
recursion).
\newc{By amalgamating  functional and concurrent constructs, 
\HOp may specify %reciprocal exchanges (protocols) 
complex session protocols that 
include both 
first-order communication (name passing)
and 
higher-order  processes (process passing)
and that
 can be 
 type-checked 
 using \emph{session types}~\cite{honda.vasconcelos.kubo:language-primitives}.
 By 
enforcing \emph{shared} and \emph{linear} usage policies,
 session types ensure that each communication channel in a
 process specification conforms to its prescribed protocol. 
Distinguishing between linear and shared channels/names is important, for session-based concurrency can be seen as involving two distinct phases:
the first one is non-deterministic and uses shared names, as it represents the interaction of processes seeking compatible protocol partners;
the second phase proceeds deterministically along linear names, as it specifies the concurrent execution of the session protocols established in the first phase.} 


%These calculi allow us to specify   
%session protocols in which higher-order values 
%(mobile code) can be exchanged in a type-safe manner. 
%; 
%governed by session types, 
%such protocols cleanly distinguish between 
%linear and unrestricted behaviors in 
%%directed %point-to-point 
%communications.
%in particular via  comparisons with the first-order mobility of the $\pi$-calculus~\cite{MilnerR:calmp1}. 
Although models of higher-order concurrency with session 
communication % higher-order features 
have been already developed (cf. works by Mostrous and Yoshida~\cite{tlca07} and by Gay and Vasconcelos~\cite{DBLP:journals/jfp/GayV10}),
their \emph{behavioural equivalences} 
remain little understood. 
Clarifying the status of these equivalences is essential to, e.g., 
justify non-trivial optimisations in protocols involving both name and process passing.
\newc{
An important aspect in the development of these typed equivalences is that typed semantics are usually {\em coarser} than untyped semantics. 
Indeed, since (session) types limit the contexts (environments) in which processes can interact, 
typed equivalences admit stronger properties than their untyped counterpart.
}
%for higher-order session calculi. 
%these two issues 
%have been thoroughly studied
%%are well-understood 
%for higher-order languages without sessions \cite{},
%but not for higher-order process calculi with sessions.
%This is unfortunate, given the wide applicability of session-based concurrency; indeed,
%session types are expressive enough to describe complex 
%communication structures found in practical protocols,  expressible, e.g., via recursive session types.
%Clarifying the status of typed equivalences and relative expressiveness for session languages

%but also for transferring key reasoning techniques between (higher-order) session calculi. 
%Our discovery is that \emph{linearity} of session types plays a vital role to 
%offer equalities/characterisations
%% and fully abstract encodability, 
%which to our knowledge have not been proposed before.   


%In this paper we study
%%address  behavioural equivalences for 
%\HOp, 
%%study behavioral equalities for \HOp, 
%an extension of the higher-order $\pi$-calculus~\cite{SangiorgiD:expmpa} with session primitives:
%\HOp contains constructs for 
%%session establishment
%synchronisation on shared names, 
%recursion, 
% (linear) session communication (value passing and
%labelled choice),
%abstractions and applications. 
%Abstractions are functions from values to processes, 
%\jpc{denoted}
%$\lambda x.P$; applications are 
%denoted $(\lambda x.P)V$, where the value $V$ is either a name or an abstraction.
%We study two significant subcalculi of \HOp, 
%\jpc{which}
%distil higher- and first-order mobility:
%the \HO-calculus, which is \HOp without recursion and name passing, and 
%the session \sessp-calculus \jpc{(here denoted~\sessp)}, which is \HOp without abstractions and applications.  
%While \sessp is, 
%in essence, the calculus in~\cite{honda.vasconcelos.kubo:language-primitives}, 
%this paper shows that \HO  is a new core calculus 
%for higher-order session concurrency.

A well-known behavioural equivalence for higher-order processes
is \emph{context bisimilarity}~\cite{San96H}. This 
 characterisation of %reduction-closed, 
barbed congruence 
offers an adequate distinguishing power at the price of heavy universal quantifications in output clauses.
Obtaining alternative 
characterisations of context bisimilarity
is thus a recurring, important problem 
for higher-order calculi---see, e.g.,~\cite{SangiorgiD:expmpa,San96H,JeffreyR05,DBLP:journals/cl/KoutavasH12,DBLP:journals/corr/Xu13a,lenglet_et_al:LIPIcs:2015:5364}. 
In particular, Sangiorgi~\cite{SangiorgiD:expmpa,San96H} has 
given %important 
%useful
characterisations of context bisimilarity
for higher-order processes; such 
characterisations, however,  %in~\cite{SangiorgiD:expmpa,San96H} 
do not scale to  
  calculi with \emph{recursive types}, which %in our experience 
  are essential to %the practice of 
  express practical protocols in 
session-based concurrency. A characterisation  
%context bisimilarity 
that solves this limitation was developed by Jeffrey and Rathke in~\cite{JeffreyR05};
their solution, however, does not consider \emph{linearity} which, as explained above, is an important aspect in session-based concurrency.

%\smallskip

\paragraph{This Work.}
Building upon~\cite{SangiorgiD:expmpa,San96H,JeffreyR05}, 
our discovery is that {linearity} as induced by session types plays a vital r\^{o}le 
%to 
%offer equalities and characterisations
% and fully abstract encodability, 
%which to our knowledge have not been proposed before. 
% 
in 
solving 
the %long-standing, 
open problem 
of characterising context bisimilarity for higher-order mobile processes with session communication.
Our approach is to exploit 
%protocol specifications given by session types to limit 
the coarser semantics induced by session types to limit
the behaviour of higher-order session processes. 
Indeed, 
the use of session typed contexts (i.e., environments disciplined by session types) 
leads to process semantics 
that admit stronger properties than untyped semantics.
 Formally, we enforce this limitation in behaviour by defining
a \emph{refined} labelled transition system (LTS)
which effectively 
narrows down the spectrum of allowed process behaviours, 
exploiting elementary processes inhabiting session types.
%thus enabling tractable reasoning techniques. 
We then introduce \emph{characteristic bisimilarity}: this  
 new notion of typed bisimilarity   is 
\emph{more tractable} than context bisimilarity, in that 
it relies on the refined LTS for input actions and, more importantly, 
does not appeal to universal quantifications on output actions. 
%shown to coincide with context bisimilarity.

Our main result is that characteristic  %tractable
bisimilarity coincides with context bisimilarity.
Besides confirming the value of characteristic bisimilarity as a useful reasoning technique for 
higher-order processes with sessions,
%for  specifications of trivial practical scenarios, 
this result is 
%also technically 
remarkable 
also from a technical perspective, for associated 
completeness proofs do not require 
operators for 
name matching,
% in the process language, 
in contrast to Jeffrey and Rathke's technique for higher-order processes
with recursive types~\cite{JeffreyR05}.
%Remarkably session type structures enable to provide 
%a coincidence without name-matching operators in the calculi.



%\smallskip

\paragraph{Outline.} 
%This paper  is structured as follows.
%\begin{enumerate}[$\bullet$]
%\item 
Next,
%%\secref{sec:overview} overviews 
we  informally overview the
key ideas of characteristic bisimilarity, 
our 
characterisation of contextual equivalence.
%%\item 
Then, \secref{sec:calculus}  presents 
%we present
the %higher-order 
session calculus \HOp. 
%A small example is given in \S\,X.
\secref{sec:types} gives the session type system for \HOp
and states type soundness.
%\item 
\secref{sec:behavioural} 
develops %\emph{higher-order} and 
\emph{characteristic} bisimilarity and 
%which alleviates the issues of context bisimilarity~\cite{San96H} and 
states our main result: characteristic bisimilarity and contextual equivalence coincide for 
%is shown  to coincide for 
well-typed \HOp processes (\thmref{the:coincidence}).
\secref{sec:relwork} discusses related works, while
%The appendix summarises the typing system. 
%\end{enumerate}
%\noi
%The paper is self-contained. 
\secref{sec:concl}~collects some concluding remarks. 

This paper is a revised, extended version of the conference paper~\cite{kouzapas_et_al:LIPIcs:2015:5365}.
This presentation includes full technical details---definitions and proofs, collected in Appendices \ref{app:types} and \ref{app:beh}.
In particular, we introduce \emph{higher-order bisimilarity} (an auxiliary labelled 
bisimilarity) 
and highlight its r\^{o}le
in the proof of \thmref{the:coincidence}. 
We also elaborate further  on the 
use case scenario 
for characteristic bisimilarity 
given in~\cite{kouzapas_et_al:LIPIcs:2015:5365} (the Hotel Booking scenario).
Using an additional example, given in  
\secref{sec:relwork}, we compare our approach with Jeffrey and Rathke's~\cite{JeffreyR05}. 
Moreover, we offer extended discussions of related works.

