% !TEX root = main.tex


%\smallskip 
We explain how we exploit session types to 
define characteristic bisimilarity.
Key notions are \emph{triggered} and \emph{characteristic processes/values}.
We first informally introduce some basic notation and terminology; formal definitions will be given in \secref{sec:calculus}.


\paragraph{Preliminaries.}
\newc{The syntax of \HOp considered in this paper is given below.
We write $n$ to range over shared names $a,b,\ldots$
and
$s, {s}', \ldots$ to range over 
session (linear) names.
Also, $u, w$ denotes a name or a name variable.
Session names are sometimes called \emph{endpoints}.
We consider a notion of \emph{duality} on names, particularly 
relevant for session names: we shall write 
  $\dual{s}$ to denote the {dual} endpoint of~$s$.
	\[
		\begin{array}{rcll}
			\text{Values}~~ V,W  & \bnfis & u & \text{names (shared and linear)} \\
			& \bnfbar &  \abs{x}{P} & \text{abstractions}
			\\[2mm]
			\text{Processes}~~ P,Q & \bnfis & \bout{u}{V}{P}  \bnfbar   \binp{u}{x}{P}  & \text{output and input} \\
			& \bnfbar &	 \bsel{u}{l} P \bnfbar \bbra{u}{l_i:P_i}_{i \in I}~~  & \text{labelled choice} \\
			& \bnfbar & \rvar{X} \bnfbar \recp{X}{P} & \text{recursion} \\
			& \bnfbar & \appl{V}{W} & \text{value application} \\
			& \bnfbar & P\Par Q \bnfbar \news{n} P \bnfbar \inact  & \text{composition, restriction, inaction}
			%\\[2mm]
		\end{array}
	\] 
}
Hence, the higher-order character of \HOp comes from the fact that $V$ can be an abstraction.

%this way, e.g., 
%$\bout{\dual{s}}{V} Q$ denotes for an output of value $V$ along endpoint 
The semantics of \HOp can be given in terms of a labelled transition system (LTS),
denoted $P \by{\ell} P'$, where $\ell$ denotes a transition label or the internal action $\tau$.
This way, for instance, 
$P \by{\bactinp{n}{V}} P'$ denotes an input transition (a value $V$ received along $n$)
and
$P \by{\news{\widetilde{m}} \bactout{n}{V}} P'$
denotes an output transition (a value $V$ emitted along $n$, extruding names $\widetilde{m}$). 
Weak transitions, written 
$P \By{\ell} P'$, abstract from internal actions in the usual way.
Throughout the paper, we write $\Re, \Re',\ldots$ to denote binary relations on (typed) processes.

\newc{
\HOp processes specify structured communications (protocols) as 
disciplined by \emph{session types}, denoted $S, S', \ldots$, 
which we informally describe next:
		\[
		\begin{array}{lcll}
		S & \bnfis &  \btout{U} S \!\! \bnfbar \!\! \btinp{U} S &  \text{output/input value of type $U$, continue as $S$}
			\\[1mm]
			&\bnfbar& \btsel{l_i:S_i}_{i \in I} \!\! \bnfbar \!\!\btbra{l_i:S_i}_{i \in I}~  & \text{internal/external labelled choice of an $S_i$}
			\\[1mm]
			& \bnfbar & \trec{t}{S} \bnfbar \vart{t} & \text{recursive protocol}
			\\[1mm]
			&\bnfbar&  \tinact  & \text{completed protocol}
		\end{array}
		\]
As we will see, type $U$ denotes first-order values (i.e., shared and session names)
but also shared and linear functional types, denoted $\shot{U}$ and $\lhot{U}$, respectively,
where $\Proc$ is the type for processes.}


%\smallskip
\paragraph{Issues of Context Bisimilarity.}
Context bisimilarity (denoted $\wbc$, cf. \defref{def:wbc}) is an overly demanding relation on higher-order processes. 
It is far from satisfactory due to two issues, associated to demanding clauses for output and input actions. 
%The first demanding issue arises from the universal quantification on the output clause of context bisimilarity:
A \emph{first issue} %appears in 
is 
the universal quantification in the output clause of context bisimilarity.
Suppose $P \,\Re\, Q$, for some context bisimulation~$\Re$. We have the following clause:
\begin{enumerate}[$(\star)$]
	\item	Whenever 
		$P \by{\news{\widetilde{m_1}} \bactout{n}{V}} P'$
		there exist
		$Q'$, $W$
		such that 
		$Q \By{\news{\widetilde{m_2}} \bactout{n}{W}} Q'$
		and, \\ \emph{\textbf{for all} $R$}  with $\fv{R}=\{x\}$, 
		$\newsp{\widetilde{m_1}}{P' \Par R\subst{V}{x}} \,\Re\, \newsp{\widetilde{m_2}}{Q' \Par R\subst{W}{x}}$.
\end{enumerate}
Intuitively, process $R$ above stands for any possible \emph{context} to which the emitted value
($V$ and $W$) is supposed to go. 
(As usual, $R\subst{V}{x}$ denotes 
the capture-avoiding substitution of $V$ for $x$ in process $R$.)
As explained in~\cite{San96H}, considering 
all possible contexts $R$ is key to achieve an adequate distinguishing power.

The \emph{second issue} is due to inputs, and  
%appears in input clauses, and 
follows from 
the fact that we work with an \emph{early}
labelled transition system (LTS). Thus, %a higher-order 
an input prefix may observe
infinitely many different values.

To alleviate these issues,
%induced by 
%universal quantification,
in %our notion of
\emph{characteristic
bisimilarity} (denoted $\fwb$, cf. \defref{d:fwb})
we take two (related) steps: 
\begin{enumerate}[(a)]
	\item We replace $(\star)$ with a clause involving a context \emph{more tractable} than $R\subst{V}{x}$ (and $R\subst{W}{x}$); and 
	\item We refine  inputs % transitions 
	to avoid observing infinitely many actions on the same input prefix.
\end{enumerate}


\begin{comment}
\myparagraph{Issues of Context Bisimilarity.}
%The characterisation of contextual congruence given by 
Context bisimilarity ($\wbc$, \defref{def:wbc}) is an overly demanding relation on higher-order processes. 
%In the following we motivate our
%proposal for alternative, more tractable characterisations.  
%For the sake of clarity, and to emphasise the novelties of our approach, 
%we often omit type information. 
%Formal definitions including types are in \S\,\ref{sec:behavioural}.
This issue may be better appreciated by examining the output clause of context bisimilarity.
% see the issue, we show its output clause.
Suppose $P \,\Re\, Q$, for some context bisimulation $\Re$. Then:
\begin{enumerate}[$(\star)$]
	\item	Whenever 
		$P \by{\news{\widetilde{m_1}} \bactout{n}{V}} P'$
		there exist
		$Q'$ and $W$
		such that 
		$Q \by{\news{\widetilde{m_2}} \bactout{n}{W}} Q'$
		and, \emph{\textbf{for all} $R$}  with $\fv{R}=x$, 
		$\newsp{\widetilde{m_1}}{P' \Par R\subst{V}{x}} \,\Re\, \newsp{\widetilde{m_2}}{Q' \Par R\subst{W}{x}}$.
\end{enumerate}
\smallskip 
\noi 
Above, 
$\news{\widetilde{m_1}} \bactout{n}{V}$ is the output label of 
value $V$ with extrusion of names in $\widetilde{m_1}$.
To reduce the burden induced by 
universal quantification, we introduce 
%\emph{higher-order}  and 
\emph{characteristic}  
bisimilarity ($\fwb$). %, a tractable equivalence.
As we work with an \emph{early} labelled transition system (LTS), 
%we shall also aim at limiting the input actions,  
%so to define a
%bisimulation relation for the output clause without observing
%infinitely many actions on the same input prefix. 
%To this end, 
%
we take the following two steps: 
%
\begin{enumerate}[(a)]
	\item replace $(\star)$ with a clause involving a more tractable process closure; and 
	\item refine the input LTS rule 
	to avoid observing infinitely many actions on the same input prefix.
\end{enumerate}
%
%\smallskip
\end{comment}
%\noi
%Before 
%explaining how we address (a) and (b), we
%introduce some  notation. We  


\paragraph{Trigger Processes.} % with Session Communication.}
To address~(a), we exploit session types. 
We first 
observe that, \newc{for any $V$}, %closure 
process $R\subst{V}{x}$ 
in $(\star)$
is context bisimilar to the process
%\begin{equation}\label{equ:1}
	$$P = \newsp{s}{\appl{(\abs{z}{\binp{z}{x}{R}})}{s} \Par \bout{\dual{s}}{V} \inact}$$
%\end{equation}
%\noi 
In fact, {through a name application and a synchronisation on session endpoint $s$}
we do have $P \wbc R\subst{V}{x}$:
\begin{eqnarray*}%\label{equ:1}
	P & \by{~\tau~} &  \newsp{s}{\appl{\binp{s}{x}{R}} \Par \bout{\dual{s}}{V} \inact} \\
	 & \by{~\tau~} &  R\subst{V}{x} \Par \inact
\end{eqnarray*}
where it is worth noticing that 
application and endpoint synchronisations 
%($s$ and $\dual{s}$) 
are deterministic.  

Now let us
consider process $T_{V}$ below, where $t$ is a fresh name:
\begin{equation}
	\label{equ:2}
	T_{V} = \binp{t}{x} \newsp{s}{\appl{x}{s} \Par \bout{\dual{s}}{V} \inact }  %\hotrigger{t}{x}{s}{V} 
\end{equation}
If $T_{V}$ inputs value $\abs{z}{\binp{z}{x} R}$ then
we have:
\begin{equation*}\label{equ:2a}
	T_{V}
	\by{\bactinp{t}{\abs{z}{\binp{z}{x} R}}} 
	R\subst{V}{x}
	\wbc 
	P
\end{equation*}
Processes such as $T_{V}$ 
offer a value at a fresh name; this class of 
{\bf\em trigger processes} 
already suggests a tractable formulation of 
bisimilarity without the demanding 
output 
clause $(\star)$. 
%However, 
%Trigger p
Process $T_{V}$ in~\eqref{equ:2} requires a higher-order communication along $t$.
%As we explain now, we can do better than this; the key is using \emph{elementary inhabitants} of session types.
As we explain below, we can give an alternative trigger process; the key is using \emph{elementary inhabitants} of session types.
%Given a fresh name $t$, we write $\htrigger{t}{V}$ to  stand for a trigger process $T_{V}$ for value $V$.

%\smallskip

%Then we can use 
%$\newsp{\tilde{m_1}}{P_1 \Par \htrigger{t}{V_1}}$ instead 
%of Clause 1) in Definition \ref{def:wbc} if we input 
%$\abs{z}{\binp{z}{x} R}$.   

\paragraph{Characteristic Processes and Values.}
\dkmargin{check narrative here---\newc{JP: seems fine now}}
To address (b), we limit the possible 
input values (such as $\abs{z}{\binp{z}{x} R}$ above) %processes 
by exploiting session types.
The key concept is that of {\bf \emph{characteristic process/value}}
of a type,  i.e.,
%The characteristic process of a session type $S$ is the process inhabiting $S$. 
a simple process term that inhabits that type (\defref{def:char}).
%Thus, characteristic processes follow the communication structures decreed by session types.
\newc{To illustrate the key idea underlying characteristic processes, consider the session type}
$$S = \btinp{\shot{S_1}} \btout{S_2} \tinact~,$$
which abstracts a protocol that 
 first
inputs an abstraction (a function from values $S_1$ to processes), 
and then outputs a value of type $S_2$.
\newc{Let $P$ be the process $\binp{u}{x} (\bout{u}{s_2} \inact \Par \appl{x}{s_1})$, where $s_1, s_2$ are fresh names.
It can be shown that $P$ inhabits session type $S$;  
for the purposes of the behavioural theory developed in this paper,
process $P$ will serve as a kind of characteristic (representative) process for $S$ along
name $u$. 
%Characteristic processes are formally given in \defref{def:char}; as
% we will see, 
%we will consider characteristic processes that contain important information extracted from (session) types.
%For type $S$ the characteristic process we consider has the form $\binp{u}{x} (\bout{t}{u} \inact \Par \appl{x}{s_1})$ for fresh name $t$.
}
%is a \emph{characteristic process} for $S$ 
%\jpc{along $u$.}


Given a session type $S$ \newc{and a name $u$}, we write $\mapchar{S}{u} $
for the characteristic process of $S$ along   $u$.
Also, %Similarly, 
given a value type $U$ (i.e., a type for channels or abstractions), we write 
$\omapchar{U}$ to denote its \emph{characteristic value} (cf.~\defref{def:char}).
As we explain next, we use 
%the %characteristic %Precisely, we exploit  the
% characteristic value %$\lambda x.\mapchar{U}{x}$. %$\lambda x.\mapchar{U}{x}$. 
$\omapchar{U}$
 to refine input transitions.

\paragraph{Refined Input Transitions.}
To refine  input transitions, we need to observe 
an additional value, 
$\abs{{x}}{\binp{t}{y} (\appl{y}{{x}})}$, 
called the {\bf\em trigger value} (cf. \defref{d:trigger}). 
This is necessary: it turns out
that a characteristic value 
alone as the observable input 
is not enough to define a sound bisimulation (cf. \exref{ex:motivation}, page \pageref{ex:motivation}).
Intuitively, the trigger value is used
to observe/simulate application processes.
%to {\em count} the number of free higher-order variables inside 
%the receiver. 
%\jpc{See Example~\ref{ex:motivation} for further details.}

\newc{Based on the above discussion, 
%Using this rule, 
we define an alternative  LTS on typed processes, denoted  $\hby{\ell}$.
We use this refined LTS to define
%both higher-order ($\hwb$) and 
{characteristic  bisimulation}
%(Defs.~\ref{d:hbw} and~\ref{d:fwb}),
%($\fwb$, 
(\defref{d:fwb}),
in which the demanding clause~$(\star)$ is replaced with 
a more tractable output clause based on 
%trigger processes \jpc{(cf.~\eqref{equ:2})}  and 
characteristic 
trigger processes
\jpc{(cf.~\eqref{eq:4} below)}.
Key to this alternative LTS is 
the following (refined) transition rule for input actions (cf. \defref{def:rlts}) 
which, roughly speaking, given some fresh $t$,
%We write $P \by{\bactinp{n}{V}} P'$ for the input transition along $n$.
%Roughly, the refined rule for  \jpc{(higher-order)} input 
only  admits names $m$, trigger values $\abs{{x}}{\binp{t}{y} (\appl{y}{{x}})}$, and characteristic values $\omapchar{U}$:}
%\[
%\tree {
%	P \by{\bactinp{n}{V}} P' \quad 
%	V = m \vee V \scong \abs{{x}}{\binp{t}{y} (\appl{y}{{x}})}
%	\vee  V \scong \omapchar{U}  \textrm{ with } t \textrm{ fresh} 
%}{
%	P' \hby{\bactinp{n}{V}} P'
%}
%\]
$$
%\boxed
{
~~P \by{\bactinp{n}{V}} P' ~\land~ (V = m \vee V \scong \abs{{x}}{\binp{t}{y} (\appl{y}{{x}})} \vee  V \scong \omapchar{U})
  %\textrm{ with } t \textrm{ fresh}) 
  ~~\Rightarrow~~
P \hby{\bactinp{n}{V}} P'~~
}
$$
%for some fresh $t$.
Note the different notation for standard and refined transitions: $\by{\bactinp{n}{V}}$ vs. $\hby{\bactinp{n}{V}}$.
%Our refined  rule for \jpc{(higher-order)} input admits only names, trigger values, and characteristic values.

%respectively.
%We show that $\hwb$ is useful for \HOp and \HO, and that~$\fwb$ can be uniformly used in all subcalculi, including \sessp. 


\paragraph{Characteristic Triggers.}
Following the same reasoning as (\ref{equ:2}), 
we can use an alternative trigger process, called
{\bf\em characteristic trigger process} to replace clause
% (1) in Definition~\ref{def:wbc}:
($\star$). 
\newc{Given a fresh name $t$ and a value $V$ of with type 
$U$, we have}: % in \defref{def:wbc}:
\begin{equation}
	\label{eq:4}
	 {
	~~\ftrigger{t}{V}{U} \defeq \fotrigger{t}{x}{s}{U}{V}~~
	}
\end{equation}

%Note that if $U=L$, $\ftrigger{t}{V}{U}$ subsumes 
%$\htrigger{t}{V}$. 
\noi
This formulation is justified, because given $T_V$ as in~\eqref{equ:2}, we may show that
$$T_V \hby{\bactinp{t}{\omapchar{\btinp{U} \tinact}}} \wbc \fotrigger{t}{x}{s}{U}{V}$$
\noi 
\jpc{Thus, unlike process~\eqref{equ:2}, the characteristic trigger process 
in~\eqref{eq:4}
does not involve a
higher-order communication on %fresh name 
$t$.}
In contrast to previous approaches~\cite{SangiorgiD:expmpa,JeffreyR05} 
our %{trigger processes} 
 characteristic trigger processes 
do {\em not} use recursion or 
replication. This is key to preserve linearity of session endpoints.  

\newc{It is also noteworthy that \HOp lacks name matching, 
which is crucial in~\cite{JeffreyR05} to prove completeness of bisimilarity.}
\dk{The lack of matching operators is compensated here with the use of (session) types.}
\newc{Matching gives the observer the ability to test the equality of received names. In contrast, 
in our theory a process trigger embeds a name into a characteristic process so as to observe its (typed) behaviour.
Thus, equivalent processes deal with (possibly different) names that have the same (typed) behaviour.}

%Instead of matching, we use types:
%a process trigger embeds a name into a characteristic
%process so to observe its session behaviour.

%\dk{We stress-out that our calculus lacks
%a matching construct
%which is usually a crucial element to prove completeness of bisimilarity.
%Nevertheless, we use types
%%and specifically the characteristic process
%to compensate;
%%for the absence of matching, i.e.~
%instead of name matching, a process trigger embeds a name into a characteristic
%process so to observe its session behavior.}
%Notice that while Definition \ref{d:hbw} is useful for 
%\HOp and its higher-order variants,
%Definition \ref{d:fwb} is useful for first-order sub-calculi of \HOp.




%%\myparagraph{Outline}
%\subsection{Outline}
%\noi \S\,\ref{sec:calculus} presents the calculi; 
%\S\,\ref{sec:types} presents types;
%the tractable bisimulations are in \S\,\ref{sec:behavioural};
%the notion of encoding is in \S\,\ref{s:expr};
%\S\,\ref{sec:positive} and \S\,\ref{sec:negative}
%present positive and negative encodability results, resp;
%\S\,\ref{sec:extension} discusses extensions; and 
%\S\,\ref{sec:relwork} concludes with related work;
%Appendix summarises the typing system. 
%The paper is self-contained. 
%{\bf\em Omitted definitions, additional related work and full proofs can be found 
%in a technical report, available from \cite{KouzapasPY15}.} 
