% !TEX root = main.tex
\section{Higher-Order Session $\pi$-Calculi}
\label{sec:calculus}

\noindent

We use the the 
\emph{Higher-Order Session $\pi$-Calculus} (\HOp)
which is a variant of the calculus originally introduced
in~\cite{characteristic_bis}.
\HOp includes both name- and abstraction-passing 
as well as recursion; it is a subcalculus of the language studied 
in~\cite{tlca07}.
The difference between \HOp and the calculus presented
in~\cite{characteristic_bis} is the fact that \HOp does
not allow higher-order value applications. We describe
the syntax in brief.
%Following the literature~\cite{JeffreyR05},
%for simplicity of the presentation
%we concentrate on the second-order call-by-value \HOp.  
(In \secref{sec:extension} we consider extensions of 
\HOp with higher-order abstractions 
and polyadicity in name-passing/abstractions.)
%We also introduce two subcalculi of \HOp. In particular, we define the 
%core higher-order session
%calculus (\HO), which 
%%. The \HO calculus is  minimal: it 
%includes constructs for shared name synchronisation and 
%%constructs for session establish\-ment/communication and 
%(monadic) name-abstraction, but lacks name-passing and recursion.

%Although minimal, in \secref{s:expr}
%the abstraction-passing capabilities of \HOp will prove 
%expressive enough to capture key features of session communication, 
%such as delegation and recursion.

\subsection{Syntax of \HOp}
\label{subsec:syntax}

\noindent
The syntax of \HOp is defined in \figref{fig:syntax}.

\input{figures/syntax}

\myparagraph{Values}
Names $a,b,c, \dots$ (resp.~$s, \dual{s}, \dots$) 
range over shared (resp. session) names. 
Names $m, n, t, \dots$ are session or shared names.
Dual endpoints are $\dual{n}$ with
$\dual{\dual{s}} = s$ and $\dual{a} = a$.
%We define the dual operation over names $n$ as $\dual{n}$ with
%$\dual{\dual{s}} = s$ and $\dual{a} = a$.
%Intuitively, names $s$ and $\dual{s}$ are dual (two) \emph{endpoints} while 
%shared names represent shared (non-deterministic) points. 
Variables are denoted with $x, y, z, \dots$, 
and recursive variables are denoted with $\varp{X}, \varp{Y} \dots$.
An abstraction %(or higher-order value) 
$\abs{x}{P}$ is a process $P$ with name parameter $x$.
%Symbols $u, v, \dots$ range over identifiers; and  $V, W, \dots$ to denote values. 
Values $V,W$ include 
identifiers $u, v, \ldots$ %(first-order values) 
and 
abstractions $\abs{x}{P}$ (first- and higher-order values, resp.). 

\myparagraph{Terms} 
include the
$\pi$-calculus prefixes for sending and receiving values $V$.
%Process $\bout{u}{V} P$ denotes the output of value $V$
%over name $u$, with continuation $P$;
%process $\binp{u}{x} P$ denotes the input prefix on name $u$ of a value
%that 
%will substitute variable $x$ in continuation $P$. 
Recursion is expressed by $\recp{X}{P}$,
which binds the recursive variable $\varp{X}$ in process $P$.
Process 
%ny
%$\appl{x}{u}$ 
$\appl{V}{u}$ 
is the application
which substitutes name $u$ on the abstraction~$V$. 
Typing  ensures that $V$ is not a name.
Processes $\bsel{u}{l} P$ and $\bbra{u}{l_i: P_i}_{i \in I}$ are the
standard session processes for selecting and branching.
%Prefix $\bsel{u}{l} P$ selects label $l$ on name $u$ and then behaves as $P$.
%Given $i \in I$ 
%Process $\bbra{u}{l_i: P_i}_{i \in I}$ offers a choice on labels $l_i$ with
%continuation $P_i$, given that $i \in I$.
%Others are standard c
Constructs for 
inaction $\inact$,  parallel composition $P_1 \Par P_2$, and 
name restriction $\news{n} P$ are standard.
Session name restriction $\news{s} P$ simultaneously binds endpoints $s$ and $\dual{s}$ in $P$.
%A well-formed process relies on assumptions for
%guarded recursive processes.
Functions $\fv{P}$ and $\fn{P}$ denotet the sets of free 
%\jpc{recursion}
variables and names; 
and \dk{assume $V$ in $\bout{u}{V}{P}$ does not include free recursive 
variables $\rvar{X}$.}
If $\fv{P} = \emptyset$, we call $P$ {\em closed}.
%; and closed $P$ without 
%free session names a {\em program}. 

\subsection{Subcalculi of \HOp}
\label{subsec:subcalculi}
\noi
We define two subcalculi of \HOp. 
%We now define several sub-calculi of \HOp. 
\begin{enumerate}[$\bullet$]
	\item	The first subcalculus is the 
		{\em core higher-order session calculus} (denoted \HO),
		which lacks recursion and name passing; its 
		formal syntax is obtained from \figref{fig:syntax} by excluding 
		constructs in \nonhosyntax{\text{grey}}.

	\item	The second subcalculus is 
		the {\em session $\pi$-calculus} 
		(denoted $\sessp$), which 
		lacks  the
		higher-order constructs
		(i.e., abstraction passing and application), but includes recursion.

%	\item	The third sub-calculus, denoted \haskp, represents cloud Haskell:
%		\[
%			\begin{array}{rclllll}
%				V,W	& ::= &		u \bnfbar  \abs{x}{P}
%				\\
%				P,Q	& ::= &		\bout{u}{m}{P}  \bnfbar  \binp{u}{x}{P} \bnfbar
%							\bsel{u}{l} P \bnfbar \bbra{u}{l_i:P_i}_{i \in I}
%				\\[1mm]
%					& \bnfbar &	\appl{V}{V} \bnfbar P\Par Q \bnfbar \news{n} P \bnfbar \inact
%		\end{array}
%		\]
\end{enumerate}
%
Let $\CAL \in \{\HOp, \HO, \sessp\}$. We write 
$\CAL^{-\mathsf{sh}}$ for $\CAL$ without shared names
(we delete $a,b$ from $n$). 
We shall demonstrate in Section~\ref{sec:positive} that 
$\HOp$, $\HO$, and $\sessp$ have the same expressivity.


\subsection{Operational Semantics}
\label{subsec:semantics}

\input{figures/semantics}

\noindent \figref{fig:reduction} (top) defines the operational semantics 
of \HOp.
Rule $\orule{App}$ is the name application; 
Rule $\orule{Pass}$ defines a shared interaction at $n$ 
(with $\dual{n}=n$) or a session interaction;  
Rule $\orule{Sel}$ is the standard rule for labelled choice/selection;%:
%given an index set $I$, 
%a process selects label $l_j$ on name $n$ over a set of
%labels $\set{l_i}_{i \in I}$ offered by a branching 
%on the dual endpoint $\dual{n}$;
and other rules are standard $\pi$-calculus rules.
Rules for \emph{structural congruence} are defined in \figref{fig:reduction} (bottom). 
We assume the expected extension of $\scong$ to values $V$.
We write $\red^\ast$ for a multi-step reduction.
