% !TEX root = main.tex

We have provided a thorough study of the expressivity of the higher-order $\pi$-calculus with sessions, here denoted $\HOp$.
Unlink most previous works on the expressivity of 
(higher-order) process calculi, we have carried out our study in the setting of \emph{session types} for structured communications.
Types not only delineate and enable process translations; they 
inform the reasoning techniques required to reason about such translations.

Prerequisites for our developments are notions of typed behavioural equivalences and abstract notions of 
\emph{encodings} that extend usual encodability criteria with type preservation criteria. 
Builds upon well-established notions in the untyped case (in particular~\cite{DBLP:journals/iandc/Gorla10})
we introduced the notion of \emph{precise  encodings} (\defref{def:goodenc}):
to our knowledge, this is the first definition of its kind. We explored precise encodability in the most representative subcalculi of \HOp, 
namely its first- and higher-order fragments, here denoted \sessp and \HO, respectively. These two sub-calculi are known to be fairly expressive in their own. A main result of precise encodability is our encoding of \HOp into \HO (\secref{subsec:HOpi_to_HO}):
since \HO lacks name passing and recursion, this typed encoding is far from trivial, also because precise encodability imposes strict conditions, which rule out plausible encoding strategies that do not preserve (session) types (cf. \secref{sec:overview}). Our study also includes extensions of \HOp with higher-order abstractions (functions from processes to processes) and polyadicity, here denoted \HOpp and \PHOp, respectively. We extend these results to their super-calculi, therefore connecting with known higher-order session calculi~\cite{tlca07}.

Overall, 
our results cover the whole spectrum of features intrinsic to higher-order session-based concurrency:
pure process-passing (with first- and higher-order abstractions), pure name-passing, polyadicity, 
linear/shared communication, and their combination (see \figref{fig:express}). Remarkably, 
the behavioural discipline embodied by 
session types turns out to be fundamental to show that all these languages are equally expressive, up to 
rather strong behavioural correspondences. Indeed, although our encodings may be exploited in an untyped setting,
session type information is critical in the proofs of key properties for preciseness, in particular full abstraction.
